---
title: "MBNMAdose outputs: Relative effects, forest plots and rankings"
author: "Hugo Pedder"
date: "`r Sys.Date()`"
output:
  knitr:::html_vignette:
    toc: TRUE
bibliography: REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{MBNMAdose}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(MBNMAdose)
#devtools::load_all()
library(rmarkdown)
library(knitr)
library(dplyr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  include=TRUE,
  tidy.opts=list(width.cutoff=80),
  tidy=TRUE
)
```

For looking at outputs from MBNMAdose we will demonstrate using
results from an Emax MBNMA on the triptans dataset:

```{r, results="hide"}
tripnet <- mbnma.network(triptans)
trip.emax <- mbnma.run(tripnet, fun=demax(emax="rel", ed50="rel")) 
```

## Estimating relative effects

It may be of interest to calculate relative effects between different
doses of agents to obtain outputs that are more similar to those from
standard NMA. A benefit of dose-response MBNMA is that relative effects between doses that have not been explored in trials can still be compared. 

Relative effects can take the form of odds ratios, mean
differences or rate ratios, depending on the likelihood and link
function used in a model. Estimating relative effects can be
particularly helpful for dose-response functions where parameter
interpretation can be challenging (e.g. splines, fractional
polynomials).

The `get.relative()` function allows for calculation of relative effects
between any doses of agents as specified by the user. This includes
doses not available in the original dataset, as these can be estimated
via the dose-response relationship. Optional arguments allow for
relative effects to be estimated at specific effect modifier values (if
meta-regression incorporated into MBNMA model), calculation of 95%
prediction intervals rather than 95% credible intervals (the default),
and for the conversion of results from the log to the natural scale. The
resulting relative effects can also be ranked (see [Ranking] for more
details).

```{r}
# Specify treatments (agents and doses) for which to estimate relative effects
treats <- list("Placebo"=0,
               "eletriptan"= 1,
               "sumatriptan"=2,
               "almotriptan"=1)

# Print relative effects on the natural scale
rels <- get.relative(trip.emax, treatments = treats, eform=TRUE)
print(rels)

# Rank relative effects
rank(rels)
```


## Forest plots

Forest plots can be easily generated from MBNMA models using the
`plot()` method on an `"mbnma"` object. By default this will plot a
separate panel for each dose-response parameter in the model. Forest
plots can only be generated for parameters which are modelled using
relative effects and that vary by agent/class.

```{r, results="hide"}
plot(trip.emax)
```


## Ranking

Rankings can be calculated for different dose-response parameters from
MBNMA models by using `rank()` on an `"mbnma"` object. Any parameter
monitored in an MBNMA model that varies by agent/class can be ranked. A
vector of these is assigned to `params`. `lower_better` indicates
whether negative responses should be ranked as "better" (`TRUE`) or
"worse" (`FALSE`).

```{r}
ranks <- rank(trip.emax, lower_better = FALSE)
print(ranks)
summary(ranks)
```

The output is an object of `class("mbnma.rank")`, containing a list for
each ranked parameter in `params`, which consists of a summary table of
rankings and raw information on agent/class (depending on argument given
to `level`) ranking and probabilities. The summary median ranks with 95%
credible intervals can be simply displayed using `summary()`.

Histograms for ranking results can also be plotted using the `plot()`
method, which takes the raw MCMC ranking results stored in `mbnma.rank`
and plots the number of MCMC iterations the parameter value for each
treatment was ranked a particular position.

```{r}
# Ranking histograms for Emax
plot(ranks, params = "emax")

# Ranking histograms for ED50
plot(ranks, params = "ed50")
```

Alternatively, cumulative ranking plots for all parameters can be
plotted simultaneously so as to be able to compare the effectiveness of
different agents on different parameters. The surface under cumulative
ranking curve (SUCRA) for each parameter can also be estimated by
setting `sucra=TRUE`.

```{r}
# Cumulative ranking plot for both dose-response parameters
cumrank(ranks, sucra=TRUE)
```

## References
