---
title: "MBNMAdose: Perform a Model-Based Network Meta-Analysis (MBNMA)"
author: "Hugo Pedder"
date: "`r Sys.Date()`"
output:
  knitr:::html_vignette:
    toc: TRUE
bibliography: REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{MBNMAdose: Perform a Model-Based Network Meta-Analysis (MBNMA)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(MBNMAdose)
#devtools::load_all()
library(rmarkdown)
library(knitr)
library(dplyr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  include=TRUE,
  tidy.opts=list(width.cutoff=80),
  tidy=TRUE
)
```

# Network meta-regression

Study-level covariates can be included in the model to adjust treatment effects following an approach for meta-regression outlined in NICE Technical Support Document 3 [@TSD3]. This can be used to explore and account for potential effect modification.

Following the definition in NICE Technical Support Document 3, network meta-regression can be expressed as an interaction on the treatment effect in arms $\geq2$:

$$\theta_{i,k}=\mu_i+(f(x,\beta_{a_{i,k}})-f(x,\beta_{a_{i,1}})) + (\psi_{1,a_{i,k}}-\psi_{1,a_{i,1}})$$

where $\theta_{i,k}$ is the linear predictor, $\mu_{i}$ is the baseline effect on arm 1, $f(x,\beta_{a_{i,k}})$ is the dose-response function at dose $x$ with dose-response parameters $\beta_{a_{i,k}}$ for agent $a$ in arm $k$ of study $i$. $\psi_{1,a_{i,k}}$ is then the effect modifying interaction between the agent in arm $k$ and the network reference agent (typically Placebo in a dose-response analysis).

## Data preparation

To improve estimation:

* continuous covariates should be centred around their mean
* binary/categorical variables should be recoded with the most commonly reported value as the reference category

```{r, reg.prep, results="hide"}
# Using the SSRI dataset
ssri.reg <- ssri

# For a continuous covariate
ssri.reg <- ssri.reg %>% 
  mutate(x.weeks = weeks - mean(weeks, na.rm=TRUE))

# For a categorical covariate
table(ssri$weeks) # Using 8 weeks as the reference
ssri.reg <- ssri.reg %>% 
  mutate(r.weeks=factor(weeks, levels=c(8,4,5,6,9,10)))

# Create network object
ssrinet <- mbnma.network(ssri.reg)
```

## Modelling

For performing network meta-regression, different assumptions can be made regarding how the effect modification may be shared across agents:

### Independent, agent-specific interactions

The least constraining assumption available in `MBNMAdose` is to assume that the effect modifier acts on each agent independently, and separate $\psi_{1,a_{i,k}}$ are therefore estimated for each agent in the network.

A slightly stronger assumption is to assume that agents within the same class share the same interaction effect, though classes must be specified within the dataset for this.

```{r, results="hide", message=FALSE}
# Regress for continuous weeks
# Separate effect modification for each agent vs Placebo
ssrimod.a <- mbnma.run(ssrinet, fun=dfpoly(degree=2), 
                     regress=~x.weeks, regress.effect = "agent")
```
```{r}
summary(ssrimod.a)
```

Within the output, a separate parameter (named `B.x.weeks[]`) has been estimated for each agent that corresponds to the effect of an additional week of study follow-up on the relative effect of the agent versus Placebo. Note that due to the inclusion of weeks as a continuous covariate, we are assuming a linear effect modification due to study follow-up.

### Random effect interaction

Alternatively, the effect modification for different agents versus the network reference agent can be assumed to be exchangeable/shared across the network about a common mean, $\hat{\psi}$, with a between-agent standard deviation of $\tau_\psi$:

$$\psi_{1,a_{i,k}} \sim N(\hat{\psi}, \tau^2_\psi)$$

```{r, results="hide", message=FALSE}
# Regress for continuous weeks
# Random effect modification across all agents vs Placebo
ssrimod.r <- mbnma.run(ssrinet, fun=dfpoly(degree=2), 
                     regress=~x.weeks, regress.effect = "random")
```
```{r}
summary(ssrimod.r)
```

In this case only a single regression paramter is estimated (`B.x.weeks`), which corresponds to the mean effect of an additional week of study follow-up on the relative effect of an active agent versus Placebo. A parameter is also estimated for the between-agent standard deviation,  `sd.B.x.weeks`.

### Common effect interaction

This is the strongest assumption for network meta-regression, and it implies that effect modification is common (equal) for all agents versus the network reference agent:

$$\psi_{1,a_{i,k}} =\hat{\psi}$$

```{r, results="hide", message=FALSE}
# Regress for categorical weeks
# Common effect modification across all agents vs Placebo
ssrimod.c <- mbnma.run(ssrinet, fun=dfpoly(degree=2), 
                     regress=~r.weeks, regress.effect = "common")
```
```{r}
summary(ssrimod.c)
```

In this case we have performed the network meta-regression on study follow-up (weeks) as a categorical covariate. Therefore, although only a single parameter is estimated for each effect modifying term, there is a separate term for each category of week and a linear relationship for effect modification is no longer assumed.

### Alternative assumptions

Although this is beyond the capability of `MBNMAdose`, one could envision a more complex model in which the interaction effect also varied by a dose-response relationship, rather than assuming an effect by agent/class or across the whole network. This would in principle contain fewer parameters than a fully independent interaction model (in which a separate regression covariate is estimated for each treatment in the dataset).

### Aggregation bias

Note that adjusting for aggregated patient-level covariates (e.g. mean age, % males, etc.) whilst using a non-identity link function can introduce aggregation bias. This is a form of ecological bias that biases treatment effects towards the null and is typically more severe where treatment effects are strong and where the link function is highly non-linear [@TSD3]. This can be resolved by performing a patient-level regression, but Individual Participant Data are required for this and such an analysis is outside the scope of `MBNMAdose`.


## Prediction using effect modifying covariates

Models fitted with meta-regression can also be used to make predictions for a specified set of covariate values. This includes when estimating relative effects using `get.relative()`. An additional argument `regress.vals` can be used to provide a named vector of covariate values at which to make predictions.

```{r}
# For a continuous covariate, make predictions at 5 weeks follow-up
pred <- predict(ssrimod.a, regress.vals=c("x.weeks"=5))
plot(pred)
```

Predictions are very uncertain for Sertraline, as studies only investigated this agent at 6 weeks follow-up and therefore the agent-specific effect modification is very poorly estimated.

```{r}
# For a categorical covariate, make predictions at 10 weeks follow-up
regress.p <- c("r.weeks10"=1, "r.weeks4"=0, "r.weeks5"=0, 
               "r.weeks6"=0, "r.weeks9"=0)

pred <- predict(ssrimod.c, regress.vals=regress.p)
plot(pred)
```

Note that categorical covariates are modelled as multiple binary dummy covariates, and so a value for each of these must be included. 


## References
