---
title: "MBNMAdose: Checking for consistency"
author: "Hugo Pedder"
date: "`r Sys.Date()`"
output:
  knitr:::html_vignette:
    toc: TRUE
bibliography: REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{MBNMAdose: Checking for consistency}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(MBNMAdose)
library(tidyverse)
#devtools::load_all()
library(rmarkdown)
library(knitr)
library(dplyr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  include=TRUE,
  tidy.opts=list(width.cutoff=80),
  tidy=TRUE
)
```

## Consistency Testing

When performing a MBNMA by pooling relative treatment effects, the
modelling approach assumes consistency between direct and indirect
evidence within a network. This is an incredibly useful assumption as it
allows us to improve precision on existing direct estimates, or to
estimate relative effects between treatments that have not been compared
in head-to-head trials, by making use of indirect evidence.

However, if this assumption does not hold, this is extremely problematic
for inference, so it is important to be able to test it. A number of
different approaches exist to allow for this in standard Network
Meta-Analysis (NMA) [@dias2013], but within dose-response MBNMA there is
added complexity because the consistency assumption can be
conceptualised either for each treatment comparison (combination of dose
and agent), or for each agent, where consistency is required for the
agent-level parameters governing the dose-response relationship. A
detailed description of consistency testing in MBNMA models is given in
Pedder et al. [-@pedder2021cons].

Testing for consistency at the agent-level is challenging as there is
unlikely to be the required data available to be able to do this -
included studies in the dataset must have multiple doses of multiple
agents, so that sufficient information is available to estimate
dose-response parameters *within that study*. However, testing for
consistency at the treatment-level is possible in MBNMA, and this is
described below. In practice, testing for consistency at the
treatment-level should suffice, as any inconsistency identified at the
treatment level should also translate to inconsistency at the agent
level and vice versa [@pedder2021cons].

Consistency also depends on the functional form assumed for the
dose-response relationship, and so is inextricably linked to model fit
of the dose-response relationship. A thorough assessment of the validity
of the fitted model is therefore important to be confident that the
resulting treatment effect estimates provide a firm basis for decision
making.

When meta-analysing dose-response studies, the potential for
inconsistency testing may actually be reasonably rare, as most (if not
all) trials will be multi-arm placebo-controlled. Since each study is
internally consistent (the relative effects within the trial will always
adhere to consistency relationships), there will be no closed loops of
treatments that are informed by independent sources of evidence.

### Unrelated Mean Effects (UME) model at the treatment level

To check for consistency at the treatment level using UME we fit an NMA
that does not assume consistency relationships, and that only models the
direct relative effects between each arm in a study and the study
reference treatment [@dias2013]. If the consistency assumption holds
true then the results from the UME model and the NMA will be very
similar. However, if there is a discrepancy between direct and indirect
evidence in the network, then the consistency assumption may not be
valid, and the UME results are likely differ in several ways:

-   The UME model may provide a better fit to the data, as measured by
    residual deviance
-   The between-study SD for different parameters may be lower in the
    UME model
-   Individual relative effects may differ in magnitude or (more
    severely) in direction for different treatment comparisons between
    UME and NMA models.

```{r, results="hide", warning=FALSE, message=FALSE}
# Using the alogliptin dataset
alognet <- mbnma.network(alog_pcfb)
nma <- nma.run(alognet, method="random")
ume <- nma.run(alognet, method="random", UME = TRUE)
```

```{r, echo=FALSE}
kable(data.frame(
  "Model"=c("NMA", "UME"),
  "resdev"=c(nma$jagsresult$BUGSoutput$median$totresdev,
                        ume$jagsresult$BUGSoutput$median$totresdev),
  "sd"=c(MBNMAdose:::neatCrI(nma$jagsresult$BUGSoutput$summary[rownames(nma$jagsresult$BUGSoutput$summary)=="sd", c(3,5,7)], digits = 2),
                       MBNMAdose:::neatCrI(ume$jagsresult$BUGSoutput$summary[rownames(ume$jagsresult$BUGSoutput$summary)=="sd", c(3,5,7)], digits=2))
), digits=2,
col.names=c("Model", "Residual Deviance", "Betwen-study SD"))
```

As both the residual deviances and between-study SDs are similar, this
would suggest that there is no evidence of inconsistency in this
network, though comparing results for individual comparisons from the
models can also be useful.

Deviance-deviance plots [@dias2013] can be compared between different
models to investigate the fit of different data points. This can be done
for two different models at a time using `devdev()`:

```{r, results="hide", warning=FALSE, message=FALSE}
# Compares residual deviance contributions from NMA and UME models
devdev(nma, ume, dev.type="resdev")
```

In Pedder et al. [-@pedder2021cons] we suggest comparing the following
models using these plots to evaluate consistency in MBNMA:

1.  Selected dose-response MBNMA model
2.  Treatment-level NMA model (using same common/random treatment
    effects as for the MBNMA)
3.  Treatment-level UME model (using same common/random treatment
    effects as for the MBNMA)

Differences in deviance contributions between models 1 and 2 will
highlight data points which may fit the dose-response relationship
poorly. Difference in deviance contributions between models 2 and 3 will
highlight data points which are suggestive of inconsistency.

### Node-splitting at the treatment level

Another approach for consistency checking is node-splitting. This splits
contributions for a particular treatment comparison into direct and
indirect evidence, and the two can then be compared to test their
similarity [@vanvalkenhoef2016]. Node-splitting in dose-response MBNMA
is an extension to this method, as indirect evidence contributions can
be calculated incorporating the dose-response function.
`mbnma.nodesplit()` takes similar arguments to `mbnma.run()`, and
returns an object of `class("nodesplit")`.

In addition to these, the argument `comparisons` can be used to indicate
which treatment comparisons to perform a nodesplit on. If left as `NULL`
(the default) node-splits will automatically be performed in all closed
loops of treatments in which comparisons are informed by independent
sources of evidence. This is somewhat similar to the function
`gemtc::mtc.nodesplit.comparisons()`, but uses a fixed network reference
treatment and therefore ensures differences between direct and indirect
evidence are parameterised as inconsistency rather than as heterogeneity
[@dias2013]. However, it also allows for indirect evidence to be
informed via the dose-response relationship even if there is no pathway
of evidence between the treatments, which can in fact lead to additional
potentially inconsistent loops. To incorporate indirect evidence in this
way `incldr=TRUE` can be set in `inconsistency.loops()`, the default
when using `mbnma.nodesplit()`

The complexity of the dose-response relationship fitted and the amount
of indirect evidence available will also affect the number of
comparisons on which node-splitting is possible [@pedder2021cons]. If
there is only limited indirect dose-response information for a given
comparison (e.g. only two doses available for the agents in the
comparison), then only a simpler dose-response function (e.g.
exponential) can be fitted. The values given in
`inconsistency.loops()$path` can give an indication as to the number of
doses available for each comparison. For example, `drparams 3 4` would
indicate that the indirect evidence is estimated only via the
dose-response relationship, and that within the indirect evidence there
are three doses available for estimating the dose-response of the agent
in `t1` of the comparison, and four doses available for estimating the
dose-responses of the agent in `t2` of the comparison. This means that a
three-parameter dose-response function would be the most complex
function that could be used when node-splitting this comparison.

As several models have to be run for each closed loop of treatments,
node-splitting can take some time to run, and it therefore is not shown
for the purposes of this vignette.

```{r, results="hide", warning=FALSE, message=FALSE, fig.show = "hide", eval=FALSE}
# Using the psoriasis dataset (>75% improvement in PASI score)
psoriasis$r <- psoriasis$r75
psorinet <- mbnma.network(psoriasis)

# Identify comparisons on which node-splitting is possible
splitcomps <- inconsistency.loops(psorinet$data.ab, incldr=TRUE)
print(splitcomps)

# If we want to fit an Emax dose-response function, there is insufficient
#indirect evidence in all but the first 6 comparisons
nodesplit <- mbnma.nodesplit(psorinet, fun=demax(), comparisons=splitcomps[1:6,], method="common")
```

```{r, eval=FALSE}
print(nodesplit)
```

Performing the `print()` method on an object of `class("nodesplit")`
prints a summary of the node-split results to the console, whilst the
`summary()` method will return a data frame of posterior summaries for
direct and indirect estimates for each split treatment comparison.

The nodesplit object itself is a list with results for each treatment
comparison that has been split. There is a lot of information within the
results, but the most useful (and easily interpretable) elements are:

-   `p.values` the Bayesian p-value for the posterior overlap between
    direct and indirect estimates
-   `quantiles` the median and 95%CrI of the posterior distributions for
    direct and indirect evidence, and for the difference between them.
-   `forest.plot` a forest plot that shows the median and 95% CrI for
    direct and indirect estimates
-   `density.plot` a plot that shows the posterior distributions for
    direct and indirect estimates

It is possible to generate different plots of each nodesplit comparison
using `plot()`:

```{r, eval=FALSE}
# Plot forest plots of direct, indirect and pooled (MBNMA) results for each comparison
plot(nodesplit, plot.type="forest")

# Plot posterior densities of direct and indirect results for each nodesplit comparisons
plot(nodesplit, plot.type="density")
```

## References
