---
title: "MBNMAdose: Exploring the data"
author: "Hugo Pedder"
date: "`r Sys.Date()`"
output:
  knitr:::html_vignette:
    toc: TRUE
bibliography: REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{MBNMAdose: Exploring the data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(MBNMAdose)
library(tidyverse)
#devtools::load_all()
library(rmarkdown)
library(knitr)
library(dplyr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  include=TRUE,
  tidy.opts=list(width.cutoff=80),
  tidy=TRUE
)
```

## Exploring the data

Before embarking on an analysis, the first step is to have a look at the
raw data. Two features (network connectivity and dose-response
relationship) are particularly important for MBNMA. For this we want to
get our dataset into the right format for the package. We can do this
using `mbnma.network()`.

```{r}
# Using the triptans dataset
network <- mbnma.network(triptans)
summary(network)
```

This function takes a dataset with the columns:

-   `studyID` Study identifiers
-   `agent` Agent identifiers (can be character, factor or numeric)
-   `dose` Numeric data indicating the dose of the given agent within
    the study arm
-   `class` An optional column indicating a particular class code.
    Agents with the same name/identifier must also have the same class
    code.

Depending on the type of data (and the likelihood) the following columns
are required:

-   Normal likelihood
    -   `y` Numeric data indicating the mean response for a given study
        arm
    -   `se` Numeric data indicating the standard error for a given
        study arm
-   Binomial likelihood
    -   `r` Numeric data indicating the number of responders in a given
        study arm
    -   `n` Numeric data indicating the total number of participants in
        a given study arm
-   Poisson likelihood
    -   `r` Numeric data indicating the number of events in a given
        study arm
    -   `E` Numeric data indicating the total exposure time in a given
        study arm

It then performs the following checks on the data:

-   The dataset has the required column names
-   There are no missing values
-   All doses are positive
-   All SE, r, N and E are positive
-   Class labels are consistent within each agent
-   Studies have at least two arms
-   Studies do not only compare the same agent at the same dose

Finally it converts the data frame into an object of
`class("mbnma.network")`, which contains indices for study arms, numeric
variables for treatments, agents and classes, and stores a vector of
treatment, agent and class names as an element within the object. By
convention, agents are numbered alphabetically, though if the original
data for agents is provided as a factor then the factor codes will be
used. This then contains all the necessary information for subsequent
`MBNMAdose` functions.

### Examining network connectivity

Examining how the evidence in the network is connected and identifying
which studies compare which treatments/agents helps to understand which
effects can be estimated, what information will be helping to inform
those estimates, and whether linking via the dose-response relationship
is possible if the network is disconnected at the treatment-level. The
complexity of dose-response relationships that can be estimated is
dependent on the number of doses of each agent available, so this is
also important to know.

Network plots can be plotted which shows which treatments/agents have
been compared in head-to-head trials. Typically the thickness of
connecting lines ("edges") is proportional to the number of studies that
make a particular comparison and the size of treatment nodes
("vertices") is proportional to the total number of patients in the
network who were randomised to a given treatment/agent (provided `N` is
included as a variable in the original dataset for `mbnma.network()`).

In `MBNMAdose` these plots are generated using `igraph`, and can be
plotted by calling `plot()`. The generated plots are objects of
`class("igraph")` meaning that, in addition to the options specified in
`plot()`, various `igraph` functions can subsequently be used to make
more detailed edits to them.

Within these network plots, vertices are automatically aligned in a
circle (as the default) and can be tidied by shifting the label distance
away from the nodes.

```{r, error = TRUE, purl = FALSE}
# Prepare data using the triptans dataset
tripnet <- mbnma.network(triptans)
summary(tripnet)

# Draw network plot
plot(tripnet)
```

If some vertices are not connected to the network reference treatment
through any pathway of head-to-head evidence, a warning will be given.
The nodes that are coloured white represent these disconnected vertices.

```{r, message=FALSE, warning=FALSE}
# Prepare data using the gout dataset
goutnet <- mbnma.network(gout)
summary(goutnet)
```

```{r}
plot(goutnet, label.distance = 5)
```

However, whilst at the treatment-level (specific dose of a specific
agent), many of these vertices are disconnected, at the agent level they
are connected (via different doses of the same agent), meaning that *via
the dose-response relationship* it is possible to estimate results.

```{r}
# Plot at the agent-level
plot(goutnet, level="agent", label.distance = 6)
```

One agent (RDEA) is still not connected to the network, but `MBNMAdose`
allows agents to connect via a placebo response *even if they do not
include placebo in a head-to-head trial* (see [Linking disconnected
treatments via the dose-response relationship]).

```{r}
# Plot connections to placebo via a two-parameter dose-response function (e.g. Emax)
plot(goutnet, level="agent", doselink = 2, remove.loops = TRUE, label.distance = 6)
```

It is also possible to plot a network at the treatment level but to
colour the doses by the agent that they belong to.

```{r, results="hide"}
# Colour vertices by agent
plot(goutnet, v.color = "agent", label.distance = 5)
```

Several further options exist to allow for inclusion of disconnected
treatments, such as assuming some sort of common effect among agents
within the same class. This is discussed in more detail later in the
vignette.

### Examining the dose-response relationship

In order to consider which functional forms may be appropriate for
modelling the dose-response relationship, it is useful to look at
results from a "split" network meta-analysis (NMA), in which each dose
of an agent is considered as separate and unrelated (i.e. we are not
assuming any dose-response relationship). The `nma.run()` function
performs a simple NMA, and by default it drops studies that are
disconnected at the treatment-level (since estimates for these will be
very uncertain if included).

```{r, results="hide", message=FALSE, warning=FALSE}
# Run a random effect split NMA using the alogliptin dataset
alognet <- mbnma.network(alog_pcfb)
nma.alog <- nma.run(alognet, method="random")
```

```{r}
print(nma.alog)

# Draw plot of NMA estimates plotted by dose
plot(nma.alog)
```

In the alogliptin dataset there appears to be a dose-response
relationship, and it also appears to be non-linear.

One additional use of `nma.run()` is that is can be used after fitting
an MBNMA to ensure that fitting a dose-response function is not leading
to poorer model fit than when conducting a conventional NMA. Comparing
the total residual deviance between NMA and MBNMA models is useful to
identify if introducing a dose-response relationship is leading to
poorer model fit. However, it is important to note that if treatments
are disconnected in the NMA and have been dropped (`drop.discon=TRUE`),
there will be fewer observations present in the dataset, which will
subsequently lead to lower pD and lower residual deviance, meaning that
model fit statistics from NMA and MBNMA may not be directly comparable.

## References
