---
title: "MBNMAdose: Package Overview"
author: "Hugo Pedder"
date: "`r Sys.Date()`"
output:
  knitr:::html_vignette:
    toc: TRUE
bibliography: REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{MBNMAdose: Package Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(MBNMAdose)
library(tidyverse)
#devtools::load_all()
library(rmarkdown)
library(knitr)
library(dplyr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  include=TRUE,
  tidy.opts=list(width.cutoff=80),
  tidy=TRUE
)
```

## Introduction

This vignette demonstrates how to use `MBNMAdose` to perform Model-Based
Network Meta-Analysis (MBNMA) of studies with multiple doses of
different agents by accounting for the dose-response relationship. This
can connect disconnected networks via the dose-response relationship and
the placebo response, improve precision of estimated effects and allow
interpolation/extrapolation of predicted response based on the
dose-response relationship.

Modelling the dose-response relationship also avoids the "lumping" of
different doses of an agent which is often done in Network Meta-Analysis
(NMA) and can introduce additional heterogeneity or inconsistency. All
models and analyses are implemented in a Bayesian framework, following
an extension of the standard NMA methodology presented by [@lu2004] and
are run in JAGS *(version 4.3.0 or later is required)* [-@jags]. For
full details of dose-response MBNMA methodology see Mawdsley et al.
[-@mawdsley2016]. Throughout this vignette we refer to a **treatment**
as a specific **dose** or a specific **agent**

This package has been developed alongside `MBNMAtime`, a package that
allows users to perform time-course MBNMA to incorporate multiple time
points within different studies. However, *they should not be loaded
into R at the same time* as there are a number of functions with shared
names that perform similar tasks yet are specific to dealing with either
time-course *or* dose-response data.

### Workflow within the package

Functions within `MBNMAdose` follow a clear pattern of use:

1.  Load your data into the correct format using `mbnma.network()` and explore potential relationships ([Exploring the data](dataexploration-1.html)
2.  Perform a dose-response MBNMA using `mbnma.run()` ([Performing a dose-response MBNMA](runmbnmadose-2.html). Modelling of effect modifying covariates is also possibly using [Network Meta-Regression](metaregression-6.html).
3.  Test for consistency at the treatment-level using functions like
    `nma.nodesplit()` and `nma.run()` ([Checking for consistency](consistencychecking-3.html)
4.  Examine model outputs, such as relative effects, forest plots and treatment rankings ([Model outputs](outputs-4.html)
5.  Use your model to predict responses using `predict()` ([Predictions](predictions-5.html)

At each of these stages there are a number of informative plots that can
be generated to help understand the data and to make decisions regarding
model fitting.

## Datasets Included in the Package

### Triptans for migraine pain relief

`triptans` is from a systematic review of interventions for pain relief
in migraine [@thorlund2014]. The outcome is binary, and represents (as
aggregate data) the number of participants who were headache-free at 2
hours. Data are from patients who had had at least one migraine attack,
who were not lost to follow-up, and who did not violate the trial
protocol. The dataset includes 70 Randomised-Controlled Trials (RCTs),
comparing 7 triptans with placebo. Doses are standardised as relative to
a "common" dose, and in total there are 23 different treatments
(combination of dose and agent). `triptans` is a data frame in long
format (one row per arm and study), with the variables `studyID`,
`AuthorYear`, `N`, `r`, `dose` and `agent`.

```{r, echo=FALSE}
kable(head(triptans), digits=2) 
```

### Biologics for treatment of moderate-to-severe psoriasis

There are 3 psoriasis datasets from a systematic review of RCTs
comparing biologics at different doses and placebo [@warren2019]. Each
dataset contains a different binary outcome, all based on the number of
patients experiencing degrees of improvement on the Psoriasis Area and
Severity Index (PASI) measured at 12 weeks follow-up. Each dataset
contains information on the number of participants who achieved
$\geq75\%$ (`psoriasis75`), $\geq90\%$ (`psoriasis90`), or $100\%$
(`psoriasis100`).

### Selective Serotonin Reuptake Inhibitors (SSRIs) for major depression

`ssri` is from a systematic review examining the efficacy of different
doses of SSRI antidepressant drugs and placebo [@furukawa2019]. The
response to treatment is defined as a 50% reduction in depressive
symptoms after 8 weeks (4-12 week range) follow-up. The dataset includes
60 RCTs comparing 5 different SSRIs with placebo.

```{r, echo=TRUE}
kable(head(ssri), digits=2) 
```

### Interventions for Serum Uric Acid (SUA) reduction in gout

`gout` is from a systematic review of interventions for lowering Serum
Uric Acid (SUA) concentration in patients with gout *[not published
previously]*. The outcome is continuous, and aggregate data responses
correspond to the mean change from baseline in SUA in mg/dL at 2 weeks
follow-up. The dataset includes 10 Randomised-Controlled Trials (RCTs),
comparing 5 different agents, and placebo. Data for one agent (RDEA)
arises from an RCT that is not placebo-controlled, and so is not
connected to the network directly. In total there were 19 different
treatments (combination of dose and agent). `gout` is a data frame in
long format (one row per arm and study), with the variables `studyID`,
`y`, `se`, `agent` and `dose`.

```{r, echo=FALSE}
kable(head(gout), digits=2) 
```

### Interventions for pain relief in osteoarthritis

`osteopain` is from a systematic review of interventions for pain relief
in osteoarthritis, used previously in Pedder et al. [-@pedder2019]. The
outcome is continuous, and aggregate data responses correspond to the
mean WOMAC pain score at 2 weeks follow-up. The dataset includes 18
Randomised-Controlled Trials (RCTs), comparing 8 different agents with
placebo. In total there were 26 different treatments (combination of
dose and agent). The active treatments can also be grouped into 3
different classes, within which they have similar mechanisms of action.
`osteopain_2wkabs` is a data frame in long format (one row per arm and
study), with the variables `studyID`, `agent`, `dose`, `class`, `y`,
`se`, and `N`.

```{r, echo=FALSE}
kable(head(osteopain), digits=2) 
```

### Alogliptin for lowering blood glucose concentration in type II diabetes

`alog_pcfb` is from a systematic review of Randomised-Controlled Trials
(RCTs) comparing different doses of alogliptin with placebo
[@langford2016]. The systematic review was simply performed and was
intended to provide data to illustrate a statistical methodology rather
than for clinical inference. Alogliptin is a treatment aimed at reducing
blood glucose concentration in type II diabetes. The outcome is
continuous, and aggregate data responses correspond to the mean change
in HbA1c from baseline to follow-up in studies of at least 12 weeks
follow-up. The dataset includes 14 RCTs, comparing 5 different doses of
alogliptin with placebo, leading to 6 different treatments (combination
of dose and agent) within the network. `alog_pcfb` is a data frame in
long format (one row per arm and study), with the variables `studyID`,
`agent`, `dose`, `y`, `se`, and `N`.

```{r, echo=FALSE}
kable(head(alog_pcfb), digits=2) 
```


### Wound closure methods for reducing Surgical Site Infection (SSI)

`ssi_closure` is from a systematic review examining the efficacy of different wound closure methods to reduce Surgical Site Infections (SSI).  The outcome is binary and represents the number of patients who experienced a SSI. The dataset includes 129 RCTs comparing 16 different interventions in 6 classes. This dataset is primarily used to illustrate how `MBNMAdose` can be used to perform different types of network meta-analysis without dose-response information. It is in long format (one row per study arm) and includes the variables `studyID`, `Year`, `n`, `r`, `trt` and `class`.

```{r, echo=FALSE}
kable(head(ssi_closure), digits=2) 
```


## References
